Job mobility and earnings growth: diffrerent markets, different outcomes.

### 2018-09-14 draft.Rmd
Update draft.Rmd

Worker mobility is a key strategy in improving outcomes like wages and conditions. Routinely, authors show that workers make significant gains through quitting jobs or securing promotions. Despite this, capitalist countries differ by job mobility rates, specifically in their rates of voluntary job mobility. Authors typically explain these differences as a difference in the hiring and firing of workers (Brown) or a difference in the rate of job creation and destruction (OECD). This article argues that mobility rates also differ due to the rewards they offer to workers. More specifically, we propose that countries differ in the rewards tied to quits and promotions.

Institutional differences in rewarding mobility rarely appear in theoretical literature. In the *job-match* model mobility stems from a mismatch between a worker's skills, and her rewards or conditions (Jovanvich). In this way countries with low rates of job mobility, are simply countries with a good record of job matching. In the *job-search* model mobility stems from the opportunity of higher wages. Thus if labour markets have a wide dispersion of wages, workers have more opportunity to match their reservation wages, which leads to more mobility. In *Human capital* theory, job mobility depends on the type of skills available to workers, either tranferable or firm specific. Thus, mobile labour markets have a high dependence on transferable skills between firms, while less mobile markets exist among firms that rely on on firm-specific skills.

Studies exploring mobility rarely consider the impact local institutions make on the observed labour market. Typically, authors use data for a given labour market and estimate the effects of quits on earnings growth for individuals or groups. Rarely do authors compare and contrast the impact of quits in two separate markets. Beyond this, authors often ignore the impact of internal promotions on earnings growth. Quits are a form of mobility, but so are promotions who have an impact on wages.

Traditionally, authors have focused extensively on what drives workers to quit. This approach often took for granted that job mobility is of interest to workers, rewarding them under certain conditions.

>>>>

Worker mobility is a key strategy for improving outcomes like wages and conditions. Routinely, authors show that workers make significant gains through quitting or securing promotions. Despite this, capitalist countries differ in their rates of job mobility; especially in rates of voluntary mobility. Authors typically explain these differences as differences in the freedom to hire or fire workers (Brown) or a differences in the rate of job creation and destruction (OECD). This article argues that mobility rates also differ due to the rewards they offer to workers. More specifically, we propose that countries differ in the rewards tied to quits and promotions.

The aim of the article is two-fold. First, we explicitly compare internal and external mobility as a means to improve eanings growth. Second, we compare this relationship under two different institutional settings, coordinated Germany and the liberal UK (Hall and Soskice, Thelen). Both aims stem from two wrinkles often found in the job mobility literature. The first is authors' tendency to ignore internal promotions as a form of voluntary mobility. Internal promotion are the main form of mobility desired by workers, yet authors often confine mobility types to external quits. Second, is the over-reliance on single case studies. Authors exploring the impact of mobility often use single country exmaples despite the fact that european labour markets are highly divided. Our article highlights this nuance by explicitly comparing quits and promotions in two different institutional settings.

The theoretical literature has little to say on country differences in mobility rates, although certain hypotheses can be drawn out. In the *job-match* model, mobility stems from a mismatch between a worker's skills and her rewards or conditions (Jovanvich). In this way countries with low rates of job mobility, are simply countries with a good record of job matching. This would make sense in both Germany and the UK, where occupational licensing is prevalent in Germany, and less so in the UK. In the *job-search* model, mobility stems from the opportunity of higher wages. Thus if labour markets have a wide dispersion of wages, workers have more opportunity to match their reservation wages, which leads to more mobility. This too makes sense in the context of Germany and the UK. In gGermany wages inequality is far lower than in the UK, thus mobility between positions is less likely to lead to "wage-quits", when compared to the UK.  In *Human capital* theory, job mobility depends on the type of skills available to workers, either tranferable or firm specific. This approach may also apply to Germany and the UK, where German employers offer training in firm specific skills, while UK workers rely most on industry specific skills and tenure.

We offer a parallel case study of two longitudinal
